### 2017-08-15 biosketch/mehedi-nihbiosketch.tex
finished biographical sketch for the submission of paper in AMIA summit 2018

%!TEX TS-program = xelatex
\documentclass{nihbiosketch}

%------------------------------------------------------------------------------

\name{Hasan, Md Mehedi}
\eracommons{MDHASAN}
\position{Graduate Teaching Assistant of Computer Science}

\begin{document}
%------------------------------------------------------------------------------

\begin{education}
Bangladesh University of Engineering and Technology, Dhaka, Bangladesh     & B.S.         & 03/2009  & Computer Science and Engineering \\
Wayne State University, Detroit, MI     & Ph.D.         & 05/2019  & Computer Science \\
\end{education}

\section{A. Personal Statement}

\begin{statement}
I have the expertise and motivation necessary to successfully carry out the proposed research project.  I have a background in computer science, with specific expertise in health informatics, machine learning, and textual data analytics.  My research includes automatic annotation of clinical text, information extraction from electronic health records for improving the study of cancer risk and outcomes.  As a graduate reserach assistant on several university- and NIH-funded grants, I successfully collaborated with other researchers, and produced the following publications.

\begin{enumerate}

\item Hasan, M., Kotov, A., Carcone, A.I., Dong, M., Naar, S. and Hartlieb, K.B. (2016). A study of the effectiveness of machine learning methods for classification of clinical interview fragments into a large number of categories. Journal of biomedical informatics, 62, 21-31.

\item Hasan, M., Kotov, A., Mohan, A., Lu, S. and Stieg, P.M. (2016). Feedback or Research: Separating Pre-purchase from Post-purchase Consumer Reviews. In European Conference on Information Retrieval, 682-688. Springer International Publishing.

\item Kotov, A., Hasan, M., Carcone, A., Dong, M., Naar-King, S. and BroganHartlieb, K. (2015). Interpretable probabilistic latent variable models for automatic annotation of clinical text. In AMIA Annual Symposium Proceedings, 2015, 785. American Medical Informatics Association.

\end{enumerate}

\end{statement}

%------------------------------------------------------------------------------
\section{B. Positions and Honors}

\subsection*{Positions and Employment}
\begin{datetbl}
2015--2017  & Graduate Research Assistant, Department of Computer Science, Wayne State University, Detroit, MI \\
2017--2017  & Part-time Faculty, Department of Computer Science, Wayne State University, Detroit, MI \\
2017--  & Graduate Teaching Assistant, Department of Computer Science, Wayne State University, Detroit, MI \\
\end{datetbl}

\subsection*{Other Experience and Professional Memberships}
\begin{datetbl}
2010--2014 		 & System Analyst and Software Developer, KB Group UK, Dhaka, Bangladesh. \\
2015--2016 		 & Research Volunteer, Henry Ford Health Systems, Detroit, MI. \\
2016--2016		 & Software Engineer Summer Intern, Title Source Inc., Detroit, MI. \\
2017--           & Reviewer, Annual Symposium of American Medical Informatics Association. \\
2017--           & Member, Association for Computing Machinery. \\
\end{datetbl}

\subsection*{Honors}
\begin{datetbl}
2003            & Excellence in H.S.C results, Islamic Foundation, Ministry of Religious Affairs, Bangladesh. \\
2004			& University Merit Scholarship, Bangladesh University of Engineering and Technology. \\
2014            & Thomas C. Rumble Fellowship, Wayne State University, Detroit, MI. \\
\end{datetbl}

%------------------------------------------------------------------------------

\section{C. Contribution to Science}

\begin{enumerate}

Manual encoding has been used to annotate clinical text of motivational interview transcripts for many years, which is time consuming and resource intensive. However, my reserach shown that state-of-the-art machine learning methods can be used as a cheap alternate for the annotation that can facilitate researchers to establish a causal relationship between different communication strategies and desired behavioral outcomes without having to repeatedly wade through pages of interview transcripts.

\end{enumerate}

\subsection*{Complete List of Published Work in MyBibliography:} 
\url{https://www.ncbi.nlm.nih.gov/sites/myncbi/1H1Fbtu2l5kQn/bibliography/53130053/public/?sort=date}

%------------------------------------------------------------------------------

\section{D. Additional Information: Research Support and/or Scholastic Performance}

\subsection*{Ongoing Research Support}

\grantinfo{R21 DK108071}{Idalski Carcone \& Kotov (PIs)}{2016--2018}
{Automated Coding of eCoaching Exchanges to Promote Healthier Eating.}
{The goal of the study is to automate qualitative communication coding using machine-learning computer science technology.}
{Role: Graduate Research Assistant}

%------------------------------------------------------------------------------

\subsection*{Completed Research Support}

\grantinfo{Ford Motor Company}{Kotov (PI)}{2014--2015}
{Extraction of actionable insights from online automotive consumer reviews.}
{The goal of this project was to extract actionable insights from online automotive consumer reviews.}
{Role: Graduate Research Assistant}

\bigskip

\grantinfo{R21 DK100760}{Idalski Carcone (PI)}{2014--2016}
{Patient-Provider Communication to Promote Health Behavior Change in African American Adolescents.}
{The goal of this study was to identify provider communication behaviors that predict motivation for behavior change and are related to changes in weight-related health behaviors.}
{Role: Graduate Research Assistant}

\bigskip

\grantinfo{NHLBI 1U01HL097889-01}{Naar--King \& Jen (PIs)}{2014-2016}
{Provider Communication Behaviors that Predict Motivation to Change in Black Adolescents with Obesity.}
{The goal of this research was to identify communication behaviors used by weight loss counselors that mostly strongly predicted Black adolescents’ motivational statements.}
{Role: Graduate Research Assistant}

\bigskip

\grantinfo{Office of Vice President for Research at Wayne State University}{2014--2016}{}
{Electronic Health Records Information Extraction for Improving the Study of Cancer Risk and Outcomes.}
{The goal of the research was to extract information from Electronic Health Records for improving the study of cancer risk and outcomes.}
{Role: Graduate Research Assistant}

\end{document}